---
title: "Abalone Report"
subtitle: "making tables and regression analysis"
author: "Melinda Higgins"
date: "09/06/2023"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      error = TRUE,
                      message = FALSE,
                      warning = FALSE)

# read in abalone dataset
# file import dataset using readr
# library(readr)
# abalone <- 
#   read_csv("abalone.csv")

# OR read in cleaned dataset and rename dataset
load("abalone_clean.RData")
abalone <- abalone_clean
```

## Summary Statistics of Abalones' - Dimensional Measurements

A useful R package for making tables is the `arsenal` package. Learn more at:

-   <https://cran.r-project.org/web/packages/arsenal/index.html> and
-   <https://mayoverse.github.io/arsenal/>.

The key function is `tableby()`, see the vignette at <https://mayoverse.github.io/arsenal/articles/tableby.html>.

```{r results = "asis"}
# NOTE: For rmarkdown to show this table correctly in 
# the final knitted output, set the R chunk option
# above to results = "asis".

# load arsenal package
library(arsenal)

# create table of summary statistics
# overall sample no groups yet
tab1 <- 
  tableby(~ length + diameter + height, 
          data=abalone)

# use the summary() function to produce the final table summary
summary(tab1)
```

## Summary Statistics of Abalones' - Weight Measurements

```{r results = "asis"}
#library(arsenal)

# put in code here to get a table of summary stats
# for wholeWeight, shuckedWeight, visceraWeight 
# and shellWeight

tab2 <- 
  tableby(~ wholeWeight + shuckedWeight + 
            visceraWeight + shellWeight, 
          data=abalone)

summary(tab2)
```

## Abalone Dimensional Measurements by Sex - default statistical tests

Now we can add a grouping variable such as comparing these summary statistics between the 3 biological sex groups: Male, Female and Infant.

Notice that the default settings produce a p-value. The `arsenal::tableby()` function is performing an ANOVA (analysis of variance) for each of these measurements.

```{r results = "asis"}
library(arsenal)

tab3 <- 
  tableby(sex ~ length + diameter + height, 
          data = abalone)

summary(tab3,
        pfootnote = TRUE)
```

## Abalone Dimensional Measurements by Sex - change statistical test

Suppose you decide that `diameter` is skewed and really need non-parametric statistics and the Kruskall-Wallis non-parametric ANOVA test performed. We can customize the statistics - learn more at <https://mayoverse.github.io/arsenal/articles/tableby.html#change-summary-statistics-within-the-formula-1> and see more options at <https://mayoverse.github.io/arsenal/articles/tableby.html#available-function-options-1>.

You'll notice this automatically creates footnotes for each customized statistic in the output table.

```{r results = "asis"}
tab4 <- 
  tableby(sex ~ anova(length, "Nmiss", "meansd") + 
            kwt(diameter, "Nmiss", "medianq1q3") + 
            anova(height, "Nmiss", "meansd"), 
          data=abalone)

summary(tab4,
        pfootnote = TRUE)
```

## Abalone Weight Measurements by Sex

```{r results = "asis"}
#library(arsenal)

# put in code here to get a table of summary stats
# for wholeWeight, shuckedWeight, visceraWeight 
# and shellWeight

tab5 <- 
  tableby(sex ~ wholeWeight + shuckedWeight + 
            visceraWeight + shellWeight, 
          data=abalone)

summary(tab5,
        pfootnote = TRUE)
```

## Plot of Abalone Age by shuckedWeight

```{r}
library(dplyr)
abalone <- abalone %>%
  mutate(age = rings + 1.5)

library(ggplot2)

ggplot(abalone,
       aes(x=shuckedWeight, 
           y=age)) +
  geom_point() +
  # add loess smoothed line
  geom_smooth(method = lm) + 
  xlab("Shucked Weight (g)") + 
  ylab("Age (in years)") + 
  ggtitle("Abalone Age by Shucked Weight")
```

## Plot of Abalone Age by Shucked Weight - by sex

Create a plot of abalone age by shucked weight in g Show the plot by sex - either add a color by sex or a facet_wrap().

```{r}
# put your code here
# option 1 color points by sex
ggplot(abalone,
       aes(x = shuckedWeight, 
           y = age,
           color = sex)) +
  geom_point() +
  # add loess smoothed line
  geom_smooth(method = lm) + 
  xlab("Diameter (mm)") + 
  ylab("Age (in years)") + 
  ggtitle("Abalone Age by Shucked Weight")

# option 2 facet_wrap() by sex
ggplot(abalone,
       aes(x = shuckedWeight, 
           y = age)) +
  geom_point() +
  # add loess smoothed line
  geom_smooth(method = lm) + 
  xlab("Diameter (mm)") + 
  ylab("Age (in years)") + 
  ggtitle("Abalone Age by Shucked Weight") +
  facet_wrap(vars(sex))
```

## Linear Regression - Abalone Age by Shucked Weight (model 1)

```{r}
lm1 <- lm(age ~ shuckedWeight,
          data = abalone)

# get nicer output
library(broom)
knitr::kable(tidy(lm1),
             caption = "Regression of Abalone Age by Shucked Weight")
```

## Linear Regression - Abalone Age by Shucked Weight, adjusted for sex (model 2)

```{r}
lm2 <- lm(age ~ shuckedWeight + sex,
          data = abalone)

# get nicer output
library(broom)
knitr::kable(tidy(lm2),
             caption = "Regression of Abalone Age by Shucked Weight adjusted for Sex")
```

## Compare Models - piecemeal steps

```{r}
slm1 <- summary(lm1)
slm2 <- summary(lm2)
mod12 <- anova(lm1, lm2)

# pull out p-value
pval12 <- mod12$`Pr(>F)`[2]
```

The change in R2 for the 2 models is `r slm2$r.squared - slm1$r.squared` with a p-value of `r pval12`.

# Compare models - use `gtsummary` package

Model 1 output

```{r}
library(gtsummary)
lm1 %>%
  tbl_regression(intercept = TRUE)
```

Model 2 output

```{r}
lm2 %>%
  tbl_regression(intercept = TRUE)
```

Put models side by side

```{r}
t1 <-
  lm1 %>%
  tbl_regression(intercept = TRUE)
t2 <-
  lm2 %>%
  tbl_regression(intercept = TRUE)

tbl_merge <-
  tbl_merge(
    tbls = list(t1, t2),
    tab_spanner = c("**Model 1**", "**Model 2**")
  ) 
tbl_merge
```

## The `stargazer` package - works for HTML and PDF

**WARNING** This does NOT work for WORD documents. However, you can create the HTML output and then "cut-and-paste" the HTML table into WORD.

```{r results = "asis"}
library(stargazer)
stargazer(lm1, lm2,
          type = "html")

# use this code if knitting to PDF
# stargazer(lm1, lm2,
#           type = "latex")
```

## Creating APA style tables

This package makes nice output but saves each table in a separate output DOC file. This does work for WORD documents.

```{r}
library(apaTables)
apa1 <- apa.reg.table(lm1,
              table.number = 1,
              filename = "table1.doc")
apa2 <- apa.reg.table(lm2,
              table.number = 2,
              filename = "table2.doc")
apa3 <- apa.reg.table(lm1, lm2,
              table.number = 3,
              filename = "table3.doc")
```

To embed the `apaTable` output inside the Rmarkdown document, we can do the following...

Pull out the key parts of the output object. Make a nice table for the formatted output for `apa3` and then add the footnote using inline r code.

```{r}
apa3[["table_body"]] %>%
  knitr::kable()
```

`r apa3[["table_note"]]`

## Visualize Regression Coefficients

Here is an example plot of the coefficients from Model 2 (`lm2`) using the packages from `easystats` - namely the `parameters` and `see` packages.

Learn more at <https://easystats.github.io/easystats/>.

```{r}
library(easystats)
library(parameters)
p2 <- parameters::model_parameters(lm2)
  
library(see)
plot(p2)
```

Here is another example using the `sjPlot` package. Note: I also had to install/update the associated `sjstats` package. Learn more at <http://www.strengejacke.de/sjPlot/reference/plot_model.html>.

```{r}
library(sjPlot)
plot_model(lm2)
```

## Logistic Regression of `adult` by Shucked Weight and Diameter

**NOTE:** The `adult` variable is currently a "character" class variable. So, let's create a 0/1 coded variable.

```{r}
abalone <- abalone %>%
  mutate(adult01 = ifelse(
    adult == "adult", 1, 0
  ))

glm1 <- glm(adult01 ~ shuckedWeight + diameter,
            data = abalone,
            family = binomial)
```

Use `parameters` package to get model coefficients table.

```{r}
library(parameters)
model_parameters(glm1,
                 exponentiate = TRUE) %>%
  knitr::kable()
```

Another option using the `gtsummary` package.

```{r}
library(gtsummary)
tbl_regression(glm1,
               exponentiate = TRUE)
```

Compute the AUC for the model and plot the ROC curve.

```{r}
library(pROC)
# get predicted outcome
pred_glm1 <- 
  predict(glm1, 
          newdata = abalone, 
          type = "response")

# compare original adult groups
# to those predicted by model
roc(abalone$adult01 ~ pred_glm1, 
    plot = TRUE, 
    print.auc = TRUE)
```
