---
title: "Abalone Report"
author: "Melinda Higgins"
date: "2023-01-12"
output:
  powerpoint_presentation: default
  slidy_presentation: default
  html_document: default
  beamer_presentation: default
  pdf_document: default
  word_document: default
  ioslides_presentation: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

# read in abalone dataset
# file import dataset using readr
library(readr)
abalone <- 
  read_csv("abalone.csv")
```

## Summary Statistics of Abalones' - Dimensional Measurements

```{r results = "asis"}
library(arsenal)

tab1 <- 
  tableby(~ length + diameter + height, 
          data=abalone)

summary(tab1)
```

## Summary Statistics of Abalones' - Weight Measurements

```{r results = "asis"}
#library(arsenal)

# put in code here to get a table of summary stats
# for wholeWeight, shuckedWeight, visceraWeight 
# and shellWeight

tab1 <- 
  tableby(~ wholeWeight + shuckedWeight + 
            visceraWeight + shellWeight, 
          data=abalone)

summary(tab1)
```

## Abalone Dimensional Measurements by Sex

```{r results = "asis"}
library(arsenal)

tab1 <- 
  tableby(sex ~ length + diameter + height, 
          data=abalone)

summary(tab1)
```

## Abalone Weight Measurements by Sex

```{r results = "asis"}
#library(arsenal)

# put in code here to get a table of summary stats
# for wholeWeight, shuckedWeight, visceraWeight 
# and shellWeight

tab1 <- 
  tableby(sex ~ wholeWeight + shuckedWeight + 
            visceraWeight + shellWeight, 
          data=abalone)

summary(tab1)
```

## Plot of Abalone Age by wholeWeight

```{r}
library(dplyr)
abalone <- abalone %>%
  mutate(age = rings + 1.5)

library(ggplot2)

ggplot(abalone,
       aes(x=wholeWeight, 
           y=age)) +
  geom_point() +
  # add loess smoothed line
  geom_smooth() + 
  xlab("Whole Weight (g)") + 
  ylab("Age (in years)") + 
  ggtitle("Abalone Age by Whole Weight")
```

## Plot of Abalone Age by diameter - by sex - Option 1

Create a plot of abalone age by diameter in mm. Show the plot by sex - Option 1 add a color by sex or Option 2 a facet_wrap().

```{r}
# put your code here
# option 1 color points by sex
ggplot(abalone,
       aes(x = diameter, 
           y = age,
           color = sex)) +
  geom_point() +
  # add loess smoothed line
  geom_smooth() + 
  xlab("Diameter (mm)") + 
  ylab("Age (in years)") + 
  ggtitle("Abalone Age by Whole Weight")
```

## Plot of Abalone Age by diameter - by sex - Option 2

Create a plot of abalone age by diameter in mm. Show the plot by sex - Option 1 add a color by sex or Option 2 a facet_wrap().

```{r}
# option 2 facet_wrap() by sex
ggplot(abalone,
       aes(x = diameter, 
           y = age)) +
  geom_point() +
  # add loess smoothed line
  geom_smooth() + 
  xlab("Diameter (mm)") + 
  ylab("Age (in years)") + 
  ggtitle("Abalone Age by Whole Weight") +
  facet_wrap(vars(sex))
```


